Actuator line modeling of vertical-axis turbines

### 2016-08-27 reviews/reviewer2-response.md

>3. Pag4. Sect.4.  Is the force smoothing with the Gaussian function really
necessary? In P.-E. Réthoré and N. N. Sørensen, A discrete force allocation
algorithm, Wind Energy, 2012. Rethore and Sorensen showed how this spatial force
filter could be avoided with modified Rhie–Chow interpolation algorithm.

In our experience, especially with the finer LES grids, the force smoothing is
necessary. Without smoothing, the blade loading becomes quite "noisy" due to the
subsequent instability from sharper momentum source gradients. In this study we
did not investigate changes to the numerical algorithms used to solve the flow
field, but certainly future research may prove the Gaussian width can be
narrowed or eliminated if, for example, modified Rhie--Chow interpolation is
implemented in OpenFOAM.

>4. Pag6. Sect.7. Does the authors took into consideration that the ALM could be
in theory able to catch the Tip vortex effect without a correction model? i.e.
the vortex generated by the lifting of the blade line is already influencing the
angle of attack on the blade.

While tip vortices are produced by the ALM, and these do affect the local angle
of attack, we are not sure how easy it would be to create a ALM that resolves
tip effects totally accurately without any correction model. The issue may be
that since we do not enforce the no peneration BC through the blade, the tip
vortex can leave the blade too early, which causes incorrect angles of attack
close to the tip.

### 2016-08-27 reviews/reviewer2-response.md

While tip vortices are produced by the ALM, and these do affect the local angle
of attack, we are not sure how easy it would be to create a ALM that resolves
tip effects totally accurately without any correction model. The issue may be
that since we do not enforce the no peneration BC through the blade, the tip
vortex can leave the blade too early, which causes incorrect angles of attack
close to the tip.

>5. Pag7. Sect.8-9 . Add a sketch of the domain boundaries with the position of
the turbine, and the flow direction. Figure 3 is not very effective in terms of
quantitative information on the actual implementation of the code.

>6. Sect. 8-9 The details on the domain boundaries of the implemented case study
should be an opportunity for clarify if any blockage effect was present in the
reference experiments and if the same blockage conditions were adopted in the
numerical implementation.

The cross-section of the domain was set equal to match blockage, though the free
surface was not modeled. However, we did use a slip BC for the top as opposed to
a fixed tow speed condition, to get closer to the free surface condition. We
have added a small clarification regarding blockage matching to the text.

@@

>7. Pag12. Sect.9.4 "The near-wake's mean velocity field at x/D=1 is shown in
Figure 12..." should be actually Figure 13

Thank you for catching this error. The manuscript has been corrected.

### 2016-08-27 paper.tex
Fix cross-referencing error
Add comment about blockage

Both the high solidity UNH-RVAT and medium solidity RM2 turbines were modeled
using the ALM inside a Reynolds-averaged Navier--Stokes (RANS) simulation,
closed with the standard $k$--$\epsilon$ turbulence model. These rotors provide
diverse parameters, which helped evaluate the robustness of the ALM. The
simulations were performed inside a domain similar in size to that used in
\cite{Bachant2016-BR-CFD} to mimic the tow tank experiments (3.66 m wide, 2.44 m
tall, 1.52 m upstream and 2.16 m downstream), with similar velocity boundary
conditions: 1 m/s inflow, no-slip (fixed to 1 m/s) walls and bottom, and a rigid
slip condition for the top, to approximate some effects of the tow tank's free
surface. Simulations were run for a total of 6 seconds, with the latter half
used to calculate performance and wake statistics. Pressure--velocity coupling
for the momentum equation was achieved using the PISO (pressure implicit
splitting of operators) method. A slice of the mesh in the $x$--$y$ plane is
shown in Figure~\ref{fig:ALM-mesh}.

@@

The near-wake's mean velocity field at $x/D=1$ is shown in
Figure~\ref{RVAT-ALM-LES-vorticity}. Compared with the RANS ALM results, the LES
looks much more like the blade-resolved and experimental results
\cite{Bachant2015-JoT}, showing the clockwise and counterclockwise mean swirling
motion on the $-y$ and $+y$ sides of the rotor, respectively.

>>>>

Both the high solidity UNH-RVAT and medium solidity RM2 turbines were modeled
using the ALM inside a Reynolds-averaged Navier--Stokes (RANS) simulation,
closed with the standard $k$--$\epsilon$ turbulence model. These rotors provide
diverse parameters, which helped evaluate the robustness of the ALM. The
simulations were performed inside a domain similar in size to that used in
\cite{Bachant2016-BR-CFD} to approximately (including blockage) the tow tank
experiments (3.66 m wide, 2.44 m tall, 1.52 m upstream and 2.16 m downstream),
with similar velocity boundary conditions: 1 m/s inflow, no-slip (fixed to 1
m/s) walls and bottom, and a rigid slip condition for the top, to approximate
some effects of the tow tank's free surface. Simulations were run for a total of
6 seconds, with the latter half used to calculate performance and wake
statistics. Pressure--velocity coupling for the momentum equation was achieved
using the PISO (pressure implicit splitting of operators) method. A slice of the
mesh in the $x$--$y$ plane is shown in Figure~\ref{fig:ALM-mesh}.

@@

The near-wake's mean velocity field at $x/D=1$ is shown in
Figure~\ref{RVAT-ALM-LES-meancontquiv}. Compared with the RANS ALM results, the
LES looks much more like the blade-resolved and experimental results
\cite{Bachant2015-JoT}, showing the clockwise and counterclockwise mean swirling
motion on the $-y$ and $+y$ sides of the rotor, respectively.
